An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-15 article/article.tex
final version

%% -- Introduction -------------------------------------------------------------

%% - In principle "as usual".
%% - But should typically have some discussion of both _software_ and _methods_.
%% - Use \proglang{}, \pkg{}, and \code{} markup throughout the manuscript.
%% - If such markup is in (sub)section titles, a plain text version has to be
%%   added as well.
%% - All software mentioned should be properly \cite-d.
%% - All abbreviations should be introduced.
%% - Unless the expansions of abbreviations are proper names (like "Journal
%%   of Statistical Software" above) they should be in sentence case (like
%%   "generalized linear models" below).

% \section[Introduction: Count data regression in R]{Introduction: Count data regression in \proglang{R}} \label{sec:intro}
\section{Introduction} \label{sec:intro}

Simulator-based models are particularly captivating due to the
modelling freedom they provide. In essence, a simulator-based model can
describe any data generating mechanism that can be written as a finite
set of algorithmic steps. This modelling freedom comes at a cost;
performing the inference, i.e., sample or evaluate the posterior
distribution, is challenging.

Optimization Monte Carlo (OMC) proposed by~\citet{Meeds2015} was a
novel LFI approach for approximating the posterior
distribution. Afterwards, \citet{Forneron2016} provided a similar
method naming it `reverse sampler'. The central idea is turning the
stochastic data-generating mechanism into a set of
fully-parallelizable deterministic optimisation processes. In their
work, ~\citet{Ikonomov2019}, located some critical failure modes of
OMC, and proposed Robust OMC (ROMC) an alternative version of OMC with
the appropriate improvements.

In this paper, we present an extendible and parallelisable
implementation of ROMC at the \proglang{Python} package \pkg{Engine
  for Likelihood-Free inference (ELFI)}. We choose carefully the
designing principles for making the implementation extendible. As we
describe analytically in the next chapter, ROMC can be understood not
only as a specific algorithm, but mainly as a framework for obtaining
weighted samples from the posterior; it defines a sequence of
algorithmic steps, without enforcing a specific algorithm for solving
them. Therefore, a researcher may adopt ROMC's backbone approach, but
experiment with alternative methods to solve the tasks. We have
designed ROMC for facilitating such alterations. Finally, we have
tested the accuracy and the efficiency of our implementation on the
LFI examples supported by the \pkg{ELFI} package.

To the best of our knowledge, this is the first attempt to implement
the ROMC inference method to a generic LFI framework. Therefore, there
are not other implementation directly comparable to ours. For this
reason, we test the implementation against (i) an artificial example
with tractable likelihood and (ii) the second-order moving average
(MA2) example from the \pkg{ELFI} package. In the latter, we generate
data using known parameters and we compare our results with the
typical Rejection ABC approach.

% \begin{leftbar}
% The introduction is in principle ``as usual''. However, it should usually embed
% both the implemented \emph{methods} and the \emph{software} into the respective
% relevant literature. For the latter both competing and complementary software
% should be discussed (within the same software environment and beyond), bringing
% out relative (dis)advantages. All software mentioned should be properly
% \verb|\cite{}|d. (See also Appendix~\ref{app:bibtex} for more details on
% \textsc{Bib}{\TeX}.)

% For writing about software JSS requires authors to use the markup
% \verb|\proglang{}| (programming languages and large programmable systems),
% \verb|\pkg{}| (software packages), \verb|\code{}| (functions, commands,
% arguments, etc.). If there is such markup in (sub)section titles (as above), a
% plain text version has to be provided in the {\LaTeX} command as well. Below we
% also illustrate how abbrevations should be introduced and citation commands can
% be employed. See the {\LaTeX} code for more details.
% \end{leftbar}

%% -- Manuscript ---------------------------------------------------------------

%% - In principle "as usual" again.
%% - When using equations (e.g., {equation}, {eqnarray}, {align}, etc.
%%   avoid empty lines before and after the equation (which would signal a new
%%   paragraph.
%% - When describing longer chunks of code that are _not_ meant for execution
%%   (e.g., a function synopsis or list of arguments), the environment {Code}
%%   is recommended. Alternatively, a plain {verbatim} can also be used.
%%   (For executed code see the next section.)

% \begin{leftbar}
% As the synopsis above is a code listing that is not meant to be executed,
% one can use either the dedicated \verb|{Code}| environment or a simple
% \verb|{verbatim}| environment for this. Again, spaces before and after should be
% avoided.

% Finally, there might be a reference to a \verb|{table}| such as
% Table~\ref{tab:overview}. Usually, these are placed at the top of the page
% (\verb|[t!]|), centered (\verb|\centering|), with a caption below the table,
% column headers and captions in sentence style, and if possible avoiding vertical
% lines.
% \end{leftbar}

%% -- Illustrations ------------------------------------------------------------

% - Virtually all JSS manuscripts list source code along with the generated
%   output. The style files provide dedicated environments for this.
% - In R, the environments {Sinput} and {Soutput} - as produced by Sweave() or
%   or knitr using the render_sweave() hook - are used (without the need to
%   load Sweave.sty).
% - Equivalently, {CodeInput} and {CodeOutput} can be used.
% - The code input should use "the usual" command prompt in the respective
%   software system.
% - For R code, the prompt "R> " should be used with "+  " as the
%   continuation prompt.
% - Comments within the code chunks should be avoided - these should be made
%   within the regular LaTeX text.

% \begin{leftbar}
% For code input and output, the style files provide dedicated environments.
% Either the ``agnostic'' \verb|{CodeInput}| and \verb|{CodeOutput}| can be used
% or, equivalently, the environments \verb|{Sinput}| and \verb|{Soutput}| as
% produced by \fct{Sweave} or \pkg{knitr} when using the \code{render_sweave()}
% hook. Please make sure that all code is properly spaced, e.g., using
% \code{y = a + b * x} and \emph{not} \code{y=a+b*x}. Moreover, code input should
% use ``the usual'' command prompt in the respective software system. For
% \proglang{R} code, the prompt \code{"R> "} should be used with \code{"+  "} as
% the continuation prompt. Generally, comments within the code chunks should be
% avoided -- and made in the regular {\LaTeX} text instead. Finally, empty lines
% before and after code input/output should be avoided (see above).
% \end{leftbar}

% \clearpage
\section{Background}
We first give a short introduction to simulator-based models and
LFI. We then focus on the methods forming the basis of this paper; OMC
and its robust improvement, ROMC. Finally, we introduce ELFI, the
software package used for our implementation.

%\input{latex_files/ch2/2_1_implicit_models}

@@

%% -- Design Principles -----------------------------------------------------
\section{Design principles}  % better title needed

\input{latex_files/ch3/3-0-design-principles.tex}

@@

%% -- ROMC implementation ----------------------------------------------------
\section{Implemented functionalities}

In this section, we present the implementation. At~\ref{subsec:ELFI}
we introduce the reader to the \proglang{Python} package \pkg{ELFI}. At~\ref{subsec:general_design} we explain the general
design that we follow. Then, we divide the implementation into three
parts; at~\ref{subsec:training} we present the training (fitting)
part, at~\ref{subsec:inference} the inference part and
at~\ref{subsec:evaluation} the evaluation part. Finally,
at~\ref{subsec:extensibility} we describe how a user may extend ROMC
with its custom modules.

@@

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}

In this section, we test the implementation using the second-order
moving average (MA2) example, which is one of the standard models of
\pkg{ELFI}. We perform the inference using three different versions of
ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
Optimisation scheme and (iii) fitting a Neural Network as a surrogate
model. The later illustrates how to extend the implementation,
replacing part of ROMC with a user-defined component. Finally, we
measure the execution speed-up achieved by the parallelised version of
ROMC.

% \subsection{Moving average example}
% \label{subsec:ma2}
\input{latex_files/ch5/5-1-ma2.tex}

% \subsection{Execution time}
% \label{subsec:exec_time}
% \input{latex_files/ch5/5-2-execution-time.tex}

%% -- Summary/conclusions/discussion -------------------------------------------
\section{Summary and discussion} \label{sec:summary}

In this paper, we presented the implementation details we followed for
developing the LFI method ROMC at the \pkg{ELFI} package. We paid
thorough attention to two specific use-case scenarios. Firstly, we
illustrate how a user may take advantage of our ready-to-use API for
solving its LFI problem. Secondly, we focus on the scenario where a
researcher wants to intervene and alter parts of the method. Our
implementation is designed to support this as well.

There are still open challenges for the left for future research. Two
directions may enable ROMC to solve high-dimensional problems
efficiently. The first one is enabling ROMC's execution into a cluster
of computers. ROMC can be characterized as an \textit{embarrassingly
  parallel} workload; each optimization problem is an entirely
independent task. Therefore, supporting inference into a cluster of
computers can radically speed up the inference. The second one refers
to implementing the method in a framework that supports automatic
differentiation. Automatic differentiation is necessary for
efficiently solving optimisation problems, especially in
high-dimensional parametric models.

@@

% \section*{Computational details}

% \begin{leftbar}
% If necessary or useful, information about certain computational details
% such as version numbers, operating systems, or compilers could be included
% in an unnumbered section. Also, auxiliary packages (say, for visualizations,
% maps, tables, \dots) that are not cited in the main text can be credited here.
% \end{leftbar}

% The results in this paper were obtained using
% \proglang{R}~3.4.1 with the
% \pkg{MASS}~7.3.47 package. \proglang{R} itself
% and all packages used are available from the Comprehensive
% \proglang{R} Archive Network (CRAN) at
% \url{https://CRAN.R-project.org/}.

@@

% \begin{leftbar}
% All acknowledgments (note the AE spelling) should be collected in this
% unnumbered section before the references. It may contain the usual information
% about funding and feedback from colleagues/reviewers/etc. Furthermore,
% information such as relative contributions of the authors may be added here
% (if any).
% \end{leftbar}

%% -- Bibliography -------------------------------------------------------------
%% - References need to be provided in a .bib BibTeX database.
%% - All references should be made with \cite, \citet, \citep, \citealp etc.
%%   (and never hard-coded). See the FAQ for details.
%% - JSS-specific markup (\proglang, \pkg, \code) should be used in the .bib.
%% - Titles in the .bib should be in title case.
%% - DOIs should be included where available.

\bibliography{refs}

%% -- Appendix (if any) --------------------------------------------------------
%% - After the bibliography with page break.
%% - With proper section titles and _not_ just "Appendix".

% \newpage

% \begin{appendix}

% \section{More technical details} \label{app:technical}

% \begin{leftbar}
% Appendices can be included after the bibliography (with a page break). Each
% section within the appendix should have a proper section title (rather than
% just \emph{Appendix}).

% For more technical style details, please check out JSS's style FAQ at
% \url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
% which includes the following topics:
% \begin{itemize}
%   \item Title vs.\ sentence case.
%   \item Graphics formatting.
%   \item Naming conventions.
%   \item Turning JSS manuscripts into \proglang{R} package vignettes.
%   \item Trouble shooting.
%   \item Many other potentially helpful details\dots
% \end{itemize}
% \end{leftbar}

% \section[Using BibTeX]{Using \textsc{Bib}{\TeX}} \label{app:bibtex}

% \begin{leftbar}
% References need to be provided in a \textsc{Bib}{\TeX} file (\code{.bib}). All
% references should be made with \verb|\cite|, \verb|\citet|, \verb|\citep|,
% \verb|\citealp| etc.\ (and never hard-coded). This commands yield different
% formats of author-year citations and allow to include additional details (e.g.,
% pages, chapters, \dots) in brackets. In case you are not familiar with these
% commands see the JSS style FAQ for details.

% Cleaning up \textsc{Bib}{\TeX} files is a somewhat tedious task -- especially
% when acquiring the entries automatically from mixed online sources. However,
% it is important that informations are complete and presented in a consistent
% style to avoid confusions. JSS requires the following format.
% \begin{itemize}
%   \item JSS-specific markup (\verb|\proglang|, \verb|\pkg|, \verb|\code|) should
%     be used in the references.
%   \item Titles should be in title case.
%   \item Journal titles should not be abbreviated and in title case.
%   \item DOIs should be included where available.
%   \item Software should be properly cited as well. For \proglang{R} packages
%     \code{citation("pkgname")} typically provides a good starting point.
% \end{itemize}
% \end{leftbar}

% \end{appendix}

>>>>

%% -- Introduction -------------------------------------------------------------
\section{Introduction} \label{sec:intro}
\input{latex_files/ch1/1-introduction.tex}

% \clearpage
\section{Background}
\input{latex_files/ch2/2-0-intro.tex}

@@

%% -- Design Principles -----------------------------------------------------
\section{Design principles}  % better title needed
\input{latex_files/ch3/3-0-design-principles.tex}

@@

%% -- ROMC implementation --
\section{Implemented functionalities}
\input{latex_files/ch4/4-0-intro.tex}

@@

%% -- Use-case illustration ----------------------------------------------------
\section{Use-case illustration}
\input{latex_files/ch5/5-ma2.tex}

%% -- Summary/conclusions/discussion -------------------------------------------
\section{Summary and discussion} \label{sec:summary}
\input{latex_files/ch6/6-summary.tex}

@@

\section*{Computational details}

The results in this paper were obtained using \proglang{Python}~3.7.1
with the \pkg{ELFI}~0.8.2 package. The experiments have been executed
in a single machine with an Intel® Core™ i7-8750H Processor an with
Ubuntu 20.04 lts operating system.

@@

\clearpage
\bibliography{refs}

### 2022-02-15 article/latex_files/ch2/2-1-implicit-models.tex
final version

The approximation in \eqref{eq:approx_likelihood} is by no means the
only strategy to deal with the intractabilities of the likelihood
function in \eqref{eq:likelihood}. Other strategies include modelling
the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
reverse, or framing likelihood-free inference as a ratio estimation
problem \citep[see e.g.\ ][]{Blum2010, Wood2010, Papamakarios2016,
  Thomas2016, Papamakarios2019, Chen2019, Thomas2020,
  Hermans2020}. However, both OMC and robust OMC, which we introduce
next, are based on the approximation in \eqref{eq:approx_likelihood}.

>>>>

The approximation in \eqref{eq:approx_likelihood} is by no means the
only strategy to deal with the intractabilities of the likelihood
function in \eqref{eq:likelihood}. Other strategies include modelling
the (stochastic) relationship between \(\thetab\) and \(\yb\), and its
reverse, or framing likelihood-free inference as a ratio estimation
problem, see for example \citet{Blum2010, Wood2010, Papamakarios2016,
  Papamakarios2019, Chen2019, Thomas2020, Hermans2020}. However, both
OMC and robust OMC, which we introduce next, are based on the
approximation in \eqref{eq:approx_likelihood}.

### 2022-02-15 article/latex_files/ch4/4-1-training.tex
final version

The training part contains six methods.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.solve_problems(n1,
                        use_bo=False,
                        optimizer_args=None,
                        seed=None)
----------------------------------------------------------------------------                    
\end{Code}

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimiser or Bayesian optimisation. These tasks are
completed sequentially, as shown in
Figure~\ref{fig:romc_overview}. The definition of the optimisation
problems is performed by drawing \code{n1} integer numbers from a
discrete uniform distribution \(u_i \sim \mathcal{U}\{1,
2^{32}-1\}\). Each integer \(u_i\) is the seed used in \pkg{ELFI}'s
random simulator. The seeds \(u_i\) are computationally analogous to
the random variables \(\vb_i\), that were defined in the previous
chapter; setting the \code{seed} eliminates all randomness of the
simulator transforming it to a deterministic function. Setting the
argument \code{use_bo=True}, chooses the Bayesian Optimisation scheme
for obtaining \(\thetab_i^*\). In this case, apart from obtaining the
optimal points \(\thetab_i^*\), we also fit a GP as a surrogate model
\(\hat{d}_i\). In this scenario, in the following steps,
\(\hat{d}_i(\thetab)\) replaces \(d_i(\thetab)\) as the distance
function.

@@

\noindent
This function helps the user decide which threshold \(\epsilon\) to
use. It plots a histogram of the distances at the optimal point
\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in case
\code{use_bo=True}. The function accepts all keyword arguments and
forwards them to the underlying \code{matplotlib.pyplot.hist()} function; in
this way the user may customise some properties of the histogram, such
as the number of bins or the range of values.

\begin{Code}
---------------------------------- python ----------------------------------  
>>> romc.estimate_regions(eps_filter,
                          use_surrogate=None,
                          region_args=None,
                          fit_models=False,
                          fit_models_args=None,
                          eps_region=None,
                          eps_cutoff=None)
----------------------------------------------------------------------------
\end{Code}

\noindent
This method constructs a bounding box around the optimal point
\(\thetab_i^* : i = 1, 2, \ldots, n_1\) following
Algorithm~\ref{alg:region_construction}. The Hessian matrix is
approximated based on the Jacobian
\(\hess_i = \jac_i^T \jac_i\)\footnote{The Jacobian matrix is computed
  on the output of the summary statistics
  \(\nabla_{\thetab} \Phi(M_d(\thetab, \vb_i))\) at \(\thetab_i^*\)} if
gradients have been used, or as the Hessian of the GP, in case of
Bayesian Optimisation. The eigenvectors are computed using the function
\code{numpy.linalg.eig()}, and used as the axes of the
bounding box. Afterwards, the limits along the
eigenvectors are computed with
Algorithm~\ref{alg:region_construction}.

@@

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \\ \code{estimate_regions} into a single
step. The function \code{visualize_region} can be used for plotting
the bounding box around the optimal point, when the parameter space is
one- or two-dimensional. The argument \code{i} is the index of the corresponding
optimisation problem i.e.\ \(i<n_1\). \code{compute_eps} returns the
appropriate distance value \(d_{i=\kappa}^*\) where
\(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number of
accepted solutions. It can be used to automate the selection of the
threshold \(\epsilon\), e.g.\ \code{eps=romc.compute_eps(quantile=0.9)}.

>>>>

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.solve_problems(n1, use_bo=False, optimizer_args=None)
----------------------------------------------------------------------------
\end{Code}

\noindent
This method (a) draws the nuisance variables, (b) defines the
optimisation problems and (c) solves them using either a
gradient-based optimiser or the Bayesian optimisation (B0) scheme. The
three tasks are completed sequentially, as shown in
Figure~\ref{fig:romc_overview}. The definition of the optimisation
problems is performed by drawing \code{n1} integer numbers from a
discrete uniform distribution \(u_i \sim \mathcal{U}\{1,
2^{32}-1\}\). Each integer \(u_i\) is the seed used in \pkg{ELFI}'s
random simulator. The user may select the Bayesian Optimisation scheme
by setting the argument \code{use_bo=True} and pass custom arguments
to the optimizer through \code{optimize_args}.

@@

\noindent
This function helps the user decide which threshold \(\epsilon\) to
use, by plotting a histogram of the distances at the optimal point
\(d_i(\thetab_i^*) : \{i = 1, 2, \ldots, n_1\}\) or \(\hat{d}_i^*\) in
case \code{use_bo=True}. The function accepts all keyword arguments
and forwards them to the underlying \code{pyplot.hist()} function of
the package \pkg{matplotlib}. In this way the user may customise some
properties of the histogram, such as the number of bins or the range
of values.

\begin{Code}
---------------------------------- python ----------------------------------
>>> romc.estimate_regions(eps_filter, use_surrogate=None, fit_models=False)
----------------------------------------------------------------------------
\end{Code}

\noindent
This method estimates the proposal region around the optimal points,
following Algorithm~\ref{alg:region_construction}. By default, the
distance model that will be used follows the decision from the
previous step; if a gradient-based optimizer has been used, then the
real distance function \(d\) will be chosen. If BO, then then the
surrogate model \(\hat{d}\). In case, the user wants to enforce using
the original distance function \(d\) after BO, they may set
\code{use_surrogate=False}. Finally, the option \code{fit_models}
defines whether to fit local surrogate models \(\tilde{d}\) after
estimating the proposal regions.

@@

\noindent
The function \code{fit_posterior} is a syntactic sugar for applying
\code{solve_problems} and \\ \code{estimate_regions} into a single
step. The function \code{visualize_region} can be used for plotting
the bounding box around the optimal point, when the parameter space is
up to 2D. The argument \code{i} is the index of the corresponding
optimisation problem. Finally, \code{compute_eps} returns the
appropriate distance value \(d_{i=\kappa}^*\) where
\(\kappa = \lfloor quantile \cdot n \rfloor\) from the collection
\(\{ d_i^* \} \forall i = \{1, \ldots, n\}\) where \(n\) is the number
of accepted solutions. It can be used to automate the selection of the
threshold \(\epsilon\), e.g.\
\code{eps=romc.compute_eps(quantile=0.9)}.

### 2022-02-15 article/latex_files/ch4/4-2-inference.tex
final version

The inference part contains four methods.

@@

\noindent
This function computes the expectation
\(\E_{p(\thetab|\data)}[h(\thetab)]\) using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. Method \code{eval_unnorm_posterior} computes the
unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
\(Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab\); with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we perform weighted sampling from the
ROMC approximate posterior. Afterwards, we use some of \pkg{ELFI}'s
built-in tools to get a summary of the obtained samples. In Figure
\ref{fig:running_example_romc_inference}, we observe the histogram of the weighted
samples and the acceptance region of the first deterministic function
(as before) alongside the samples obtained from it. Finally, in the
code snippet, we demonstrate how to use the \code{compute_expectation}
function; in the current example, we define \code{h} to compute
firstly the empirical mean and afterwards the empirical variance. In
both cases, the empirical result is close to the ground truth
\(\mu = 0\) and \(\sigma^2 = 1\). The
\code{romc.eval_unnorm_posterior(theta)} evaluates the posterior at
point \(\theta\) using expression \eqref{eq:approx_posterior_omc}. The
\code{romc.eval_posterior(theta)} approximates the partition function
\(Z = \int_{\thetab} p_{d,\epsilon}(\thetab|\data) d\thetab\) using the
Riemann approximation as explained above. In our simple example, this
utility can provide a nice plot of the approximate posterior as
illustrated in Figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth posterior.

>>>>

@@

\noindent
The function \code{compute_expectation} computes the expectation
\(\E_{p(\thetab|\data)}[h(\thetab)]\) using
expression~\eqref{eq:expectation}. The argument \code{h} can be any
python \code{Callable}. Method \code{eval_unnorm_posterior} computes
the unnormalised posterior approximation using
expression~\eqref{eq:approx_posterior}. Method \code{eval_posterior}
evaluates the normalised posterior estimating the partition function
\(Z = \int p_{d,\epsilon}(\thetab|\data)d\thetab\); with Riemann's
integral approximation. The approximation is computationally feasible
only in low-dimensional parametric spaces.

@@

In the following code snippet, we use the inference utilities to (a)
get weighted samples from the approximate posterior, (b) compute an
expectation and (c) evaluate the approximate posterior. We also use
some of \pkg{ELFI}'s built-in tools to get a summary of the obtained
samples. For the utility \code{compute_expectation}, we demonstrate
how to use it in order to compute the samples mean and
variance. Finally, we evaluate \code{eval_posterior} at multiple
points to plot the approximate posterior of
Figure~\ref{fig:running_example_romc_inference}. We observe that the
approximation is quite close to the ground-truth.

### 2022-02-15 article/latex_files/ch5/5-ma2.tex
final version

@@

As already stated, it is straightforward for the two basic parts of
the training part to be parallelised. There is a batch of \(n_1\)
independent objectives functions to be optimised and \(n_1\) bounding
boxes to be built. Both tasks can be applied in a parallel utilising
all the available CPU cores. Our implementation supports
parallelisation using the built-in \proglang{Python} package
\pkg{multiprocess}. In Figure~\ref{fig:exec_parallel} we observe the
execution times for performing the inference. The parallel version
performs all tasks between 2.5 and 6 times faster compared to the
sequential; the optimisation problems are solved almost 6 times
faster, sampling 3.5 times faster and evaluating the posterior and
constructing the bounding boxes almost 2.5 times faster.

\begin{figure}[ht]
  \begin{center}
    \resizebox{.47\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_solve.tex}
    }
    \resizebox{.49\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_regions.tex}
    }
    \end{center}
    \caption[Execution time exploiting parallelisation]{In the first
      line, we compare the parallel and sequential execution of the
      training part and at the second line the inference part. At the
      left figure, we measure the execution time for sampling \(n_2=50\)
      points per region. At the right figure we measure the execution
      time for evaluating the posterior at a batch of \(50\) points.}
      \label{fig:exec_parallel}
\end{figure}

% \begin{figure}[ht]
%   \centering
%     \begin{minipage}[t]{.4\textwidth}
%       \begin{tikzpicture}
%         \begin{axis}[
%           title={Solve optimisation problems},
%           xlabel={\(n_1\)},
%           ylabel={seconds},
%           xmin=1, xmax=500,
%           ymin=0, ymax=7,
%           xtick={1,100,200,300,400,500},
%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
%           legend pos=north west,
%           ymajorgrids=true,
%           grid style=dashed,
%           ]
%           \addplot[color=blue, mark=o] coordinates {
%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
%           };
%           \addlegendentry{sequential}
%           \addplot[color=red, mark=o] coordinates {
%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
%           };
%           \addlegendentry{parallel}          
%         \end{axis}
%       \end{tikzpicture}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{.4\textwidth}
%       \begin{tikzpicture}
%         \begin{axis}[
%           title={Proposal regions},
%           xlabel={\(n_1\)},
%           ylabel={seconds},
%           xmin=1, xmax=500,
%           ymin=0, ymax=7,
%           xtick={1,100,200,300,400,500},
%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
%           legend pos=north west,
%           ymajorgrids=true,
%           grid style=dashed,
%           ]
%           \addplot[color=blue, mark=o] coordinates {
%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
%           };
%           \addlegendentry{sequential}
%           \addplot[color=red, mark=o] coordinates {
%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
%           };
%           \addlegendentry{parallel}          
%         \end{axis}
%       \end{tikzpicture}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{.4\textwidth}
%       \begin{tikzpicture}
%         \begin{axis}[
%           title={Sample},
%           xlabel={\(n_1\)},
%           ylabel={seconds},
%           xmin=1, xmax=500,
%           ymin=0, ymax=7,
%           xtick={1,100,200,300,400,500},
%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
%           legend pos=north west,
%           ymajorgrids=true,
%           grid style=dashed,
%           ]
%           \addplot[color=blue, mark=o] coordinates {
%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
%           };
%           \addlegendentry{sequential}
%           \addplot[color=red, mark=o] coordinates {
%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
%           };
%           \addlegendentry{parallel}          
%         \end{axis}
%       \end{tikzpicture}
%     \end{minipage}\hfill
%     \begin{minipage}[t]{.4\textwidth}
%       \begin{tikzpicture}
%         \begin{axis}[
%           title={Evaluate posterior},
%           xlabel={\(n_1\)},
%           ylabel={seconds},
%           xmin=1, xmax=500,
%           ymin=0, ymax=7,
%           xtick={1,100,200,300,400,500},
%           ytick={0, 1, 2, 3, 4, 5, 6, 7},
%           legend pos=north west,
%           ymajorgrids=true,
%           grid style=dashed,
%           ]
%           \addplot[color=blue, mark=o] coordinates {
%             (1,0.05)(125,1.6)(250,3.5)(375,5)(500,6.8)
%           };
%           \addlegendentry{sequential}
%           \addplot[color=red, mark=o] coordinates {
%             (1,0.2)(125,0.6)(250,1)(375,1.5)(500,2)
%           };
%           \addlegendentry{parallel}          
%         \end{axis}
%       \end{tikzpicture}
%     \end{minipage}\hfill
%     \caption[Execution time exploiting parallelisation]{In the first
%       line, we compare the parallel and sequential execution of the
%       training part and at the second line the inference part. At the
%       left figure, we measure the execution time for sampling \(n_2=50\)
%       points per region. At the right figure we measure the execution
%       time for evaluating the posterior at a batch of \(50\) points.}
%       \label{fig:exec_parallel}
% \end{figure}

>>>>

In this section, we test the implementation using the second-order
moving average (MA2) example, which is one of the standard models of
\pkg{ELFI}. We perform the inference using three different versions of
ROMC; (i) using a gradient-based optimiser, (ii) using the Bayesian
Optimisation scheme and (iii) fitting a Neural Network as a surrogate
model. The later illustrates how to extend the implementation,
replacing part of ROMC with a user-defined component. Finally, we
measure the execution speed-up achieved by the parallelised version of
ROMC.

@@

As stated above, ROMC is a ridiculously parallelisable
method. Therefore, it is straightforward to parallelise the fitting
part, i.e. (i) solving the optimisation problems and (ii) estimating
the proposal regions. Our implementation supports exploiting all the
available CPU cores through the built-in \proglang{Python} package
\pkg{multiprocess}\footnote{https://docs.python.org/3/library/multiprocessing.html}. In
Figure~\ref{fig:exec_parallel} we observe the execution times for
performing the inference on the ma2 model; the parallel version
performs both tasks almost six times faster than the sequential. The
result is reasonable given that the experiments have run in a single
machine with the Intel® Core™ i7-8750H Processor, which has six
separate cores.

\begin{figure}[ht]
  \begin{center}
    \resizebox{.49\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_solve.tex}
    }
    \resizebox{.49\columnwidth}{!}{%
      \input{./latex_files/images/chapter4/exec_time_regions.tex}
    }
    \end{center}
    \caption[Execution time exploiting parallelisation]{Comparison
      between parallel and sequential execution of the fitting part of
      ROMC. We observe that the parallel version runs between 2.5 an 6
      times faster.}
      \label{fig:exec_parallel}
\end{figure}
